---
title: "Covid91 vaccine study Final2023"
author: "javalos16"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalData <-read.csv(file="final.csv",header=TRUE)
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.



## Males

$H_0$; The vaccine is not effective in preventing the spread of Covid in males

$H_A$; The vaccine is effective in preventing the spread of Covid in males 

Data for this study came from a FinalData file, provided by the professor.


##Methods

The response method used is cat~cat. Data used was based on data from FinalData. We have to subset the data to include male.

Graphical display: bar chart, numerical display and inferential results


numerical display

##Results

```{r}
maleData = filter (FinalData,sex == "M")
```

##Descriptive Results

##Graphical Descriptive Results


```{r}
barchartGC(~infected+treatment,data=maleData)
barchartGC(~infected+treatment,data=maleData, type="percent")
```

The graph shows among those who have covid vaccine, the people who did get vaccine there was a difference. Meanwhile Among the people who were OK that toke the vaccine, there was no significant difference.

##Numerical Descriptive Results

```{r}
table1 <- xtabs(~infected + treatment, data=maleData)
rowPerc(table1)
colPerc(table1)
```

Among the people who toke the vaccine while they have Covid there was a 2X difference. The people that took the placebo with 70% chance of not being infected and 30% less likely for the peolple that took the vaccine. Among the people that were Ok and took the vaccine there was no significant difference. 

##Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


The Chi-squared value is .00000000000004631 is too small, so vaccine is effective in preventing spread of vaccine among drug users.

Based on the Fisher's Exact vaccine  you are 0.35 less likely to get aids. 

We will reject the null hypothesis.If you take the placebo you are 3X more likely to have aids.

## Females

$H_0$; The vaccine is not effective in preventing the spread of Covid in Females

$H_A$; The vaccine is effective in preventing the spread of Covid in Females 

Data for this study came from a FinalData file, provided by the professor.


##Methods

The response method used is cat~cat. Data used was based on data from FinalData. We have to subset the data to include female.

Graphical display: bar chart, numerical display and inferential results


numerical display


##Results

```{r}
femaleData = filter(FinalData, sex == "F")
```

##Descriptive Results

##Graphical Descriptive Results

```{r}
barchartGC(~infected+treatment,data=femaleData)
barchartGC(~infected+treatment,data=femaleData, type="percent")
```

##Numerical Descriptive Results

```{r}
table2 <- xtabs(~infected + treatment, data=femaleData)
rowPerc(table2)
colPerc(table2)
```

##Inferential Results

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

## LGBTQ

$H_0$; The vaccine is not effective in preventing the spread of Covid in LGBTQ

$H_A$; The vaccine is effective in preventing the spread of Covid in LGBTQ 

Data for this study came from a FinalData file, provided by the professor.


##Methods

The response method used is cat~cat. Data used was based on data from FinalData. We have to subset the data to include LGBTQ.

Graphical display: bar chart, numerical display and inferential results


numerical display


##Results

```{r}
LGBTQData = filter(FinalData, LGBTQ == "gay")
```

##Descriptive Results

##Graphical Descriptive Results

```{r}
barchartGC(~infected+treatment,data=LGBTQData)
barchartGC(~infected+treatment,data=LGBTQData, type="percent")
```

describe

##Numerical Descriptive Results

```{r}
table3 <- xtabs(~infected + treatment, data=LGBTQData)
rowPerc(table3)
colPerc(table3)
```

##Inferential Results

```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```

## Druggies

$H_0$; The vaccine is not effective in preventing the spread of Covid in Druggies

$H_A$; The vaccine is effective in preventing the spread of Covid in Druggies 

Data for this study came from a FinalData file, provided by the professor.


##Methods

The response method used is cat~cat. Data used was based on data from FinalData. We have to subset the data to include Druggies.

Graphical display: bar chart, numerical display and inferential results


numerical display


##Results

```{r}
DruggiesData = filter(FinalData, DrugUser == "yes")
```

##Descriptive Results

##Graphical Descriptive Results

```{r}
barchartGC(~infected+treatment,data=DruggiesData)
barchartGC(~infected+treatment,data=DruggiesData, type="percent")
```

describe

##Numerical Descriptive Results

```{r}
table4 <- xtabs(~infected + treatment, data=DruggiesData)
rowPerc(table4)
colPerc(table4)
```

describe

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```


The Chi-squared value is p-value = .000000000232 is too small, so vaccine is not effective in preventing spread of Covid among other people.

Based on the Fisher's Exact vaccine  you are 0.126 less likely to get covid. 

We will Approve the null hypothesis. The information provided shows that the vaccine does not prevent the spread of Covid.



# Overall Results and Conclusions

The original question was, is the vaccine effective in preventing the spread of Covid. Base on Data provided we have found that the vaccine is not effective in stoping the spread of Covid.

