---
title: "Covid91 vaccine study Final2023"
author: "javalos16"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`" 
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
FinalDat <-read.csv(file="final.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- FinalDat  %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

A new  preventative vaccine (covid921) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

$H_0$; The vaccine is not effective in preventing the spread of Covid.

$H_A$; The vaccine is effective in preventing the spread of Covid. 

Data for this study came from a FinalData file, provided by the professor.

## Males

```{r}
barchartGC(~infected + sex,data=FinalData)
barchartGC(~infected + sex,data=FinalData, type="percent")

```

## Females




## LGBTQ

```{r}
barchartGC(~infected + LGBTQ,data=FinalData)
barchartGC(~infected + LGBTQ,data=FinalData, type="percent")

```

## Druggies

```{r}
barchartGC(~infected + DrugUser,data=FinalData)
barchartGC(~infected + DrugUser,data=FinalData, type="percent")

```

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->


```{r}
barchartGC(~infected + treatment,data=FinalData)
barchartGC(~infected + treatment,data=FinalData, type="percent")

```

The graph shows among those who had the vaccine, the people who did get the vaccine  there was no difference. Among the people who did take the placebo, the people who did not take the vaccine,there was no significant difference.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment, data=FinalData)
rowPerc(table1)
colPerc(table1)
```

Among the people who had Covid and took the vaccine there was no significant difference. Among the people that were ok and took placebo and vaccine there was no significant difference.  



## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->


```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


The Chi-squared value is p-value = .000000000000002178 is too small, so vaccine is not effective in preventing spread of Covid among other people.

Based on the Fisher's Exact vaccine  you are 1.84 less likely to get covid. 

We will Approve the null hypothesis. The information provided shows that the vaccine does not prevent the spread of Covid.



# Overall Results and Conclusions

The original question was, is the vaccine effective in preventing the spread of Covid. Base on Data provided we have found that the vaccine is not effective in stoping the spread of Covid.

